## Homework 5

<!--
Set the working directory.
No Rchunk is needed to accomplish this.
-->



<!--
Read in the homework5.csv data. For each of the following steps you
should preface your R commands with some plain English explanation.
The data are fabricated and include the number of words per minute
that people can type when listening to no music and when listening
to heavy metal.
-->
```{r}
dat<-read.csv("homework5.csv")
```


<!--
Check the structure of the data.
-->

```{r}
str(dat)
```

<!--
Find the mean words per minute for the two groups.
-->
```{r}
mean(dat$wpm[dat$music=="metal"])
mean(dat$wpm[dat$music=="none"])
```


<!--
Run an independent t-test to determine if the mean wpm for the
no music group is different from the mean wpm from heavy metal
group. Interpret the results of your analysis after you run it.
-->
```{r}
t.test(dat$wpm[dat$music=="metal"],dat$wpm[dat$music=="none"],var.equal=TRUE)
```
t(64) = 2.33, p = .02298
These results are statistically significant and indicate that the means for the two groups are different. 

<!--
Run another independent t-test with the alternative hypothesis that
the heavy metal group types faster than the no music group. Interpret
the results of your analyses.
-->

```{r}
t.test(dat$wpm[dat$music=="metal"],dat$wpm[dat$music=="none"],var.equal=TRUE, alternative = "greater")
```
t(64) = 2.332, p = .0115
These results indicate the heavy metal music group type faster than the group with no music. 

<!--
Create a bar graph of the data.
-->

```{r}
library(ggplot2)
library(gplots)
library(dplyr)
```

```{r}
f<-ggplot(temp,aes(x=factor(music),y=means))+
  geom_bar(stat="identity", color="black",
           fill=c("deepskyblue2", "deeppink"))+
  geom_errorbar(aes(ymax=means+sems,
                    ymin=means-sems),
                width=.2)+
  ggtitle("Homework Graph")+
  labs(x="music", y="Words per minute (min)")+
  scale_x_discrete(breaks=c("music","wpm"),
                   labels=c("music","words per minute"))+
  theme(plot.title=element_text(size=15,
                                face="bold",
                                vjust=.5))+
  theme(axis.title.x=element_text(size=12,
                                  face="bold",
                                  vjust=-.25))+
  theme(axis.title.y=element_text(size=12,
                                  face="bold",
                                  vjust=.25))+
  theme(axis.text.x=element_text(size=10,
                                 face="bold",
                                 color="black"))+
  theme(axis.text.y=element_text(size=10,
                                 face="bold",
                                 color="black"))+
  coord_cartesian(ylim=c(min(temp$means)-2*max(temp$sems),
                         max(temp$means)+2*max(temp$sems)))+
  theme(panel.border=element_blank(),
        axis.line=element_line())+
  theme(panel.grid.major.x=element_blank())+
  theme(panel.grid.major.y=element_line(color="darkgrey"))+
  theme(panel.grid.minor.y=element_blank())
f
```

<!--
Conduct and interpret a two-tailed paired t-test on the data.
-->
```{r}
t.test(dat$wpm[dat$music=="metal"],
        dat$wpm[dat$music=="none"],
        var.equal=TRUE,
        paired=TRUE)
```
The paired t-test indicates that the participants in the metal music condition typed faster than the participants in the condition with no music,t(64) = 12.2, p < .05
